\documentclass[12pt]{article}
% Page setup
\usepackage[margin=1in]{geometry}
\usepackage{parskip}
\usepackage{titling}
\setlength{\droptitle}{-6em} 

% Math packages
\usepackage{amsmath, amsthm, amssymb, amsfonts}
\usepackage{mathtools}
\makeatletter
\g@addto@macro\th@plain{\normalfont} % remove italics from theorems (plain style)
\makeatother

\usepackage{hyperref}
\hypersetup{hidelinks}

% Commutative diagrams and proofs
\usepackage{tikz-cd}
\usepackage{proof}
\usepackage{stmaryrd}

% Unordered lists
\usepackage{enumitem}
\newlist{related}{itemize}{1}
\setlist[related]{label=\textbullet}
\newlist{distinct}{itemize}{1}
\setlist[distinct]{label=--}


% Theorem environments
\newtheorem{theorem}{Theorem}[subsection]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{fact}[theorem]{Fact}
\newtheorem{remark}[theorem]{Remark}
\newtheorem*{remark*}{Remark}
\newtheorem*{definition*}{Definition}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}

% Useful commands for category theory
\newcommand{\alp}{\mathbf{Alph}}
\newcommand{\mc}{\mathcal}
\newcommand{\ini}{\operatorname{ini}}
\newcommand{\ter}{\operatorname{ter}}
\newcommand{\cod}{\operatorname{cod}}
\newcommand{\dom}{\operatorname{dom}}

\newcommand{\List}{\operatorname{List}}
\newcommand{\lop}{\operatorname{[ \ ]}}
\newcommand{\len}{\operatorname{len}}
\newcommand{\type}{\operatorname{type}}

\newcommand{\Exc}{\operatorname{Exception}}

\newcommand{\Kl}{\operatorname{Kl}}
\newcommand{\plus}{\operatorname{+}}

\newcommand{\Cat}{\mathcal{C}at}
\newcommand{\Vect}{\mathrm{Vect}_\mathbf{k}}
\newcommand{\Hom}{\operatorname{Hom}}
\newcommand{\Nat}{\operatorname{Nat}}
\newcommand{\id}{\mathrm{id}}
\newcommand{\op}{\mathrm{op}}
\newcommand{\obj}{\operatorname{obj}}
\newcommand{\arr}{\operatorname{arr}}
\newcommand{\adj}{\operatorname{adj}}
\newcommand{\Par}{\operatorname{par}}

\newcommand{\coadj}{\operatorname{coadj}}
\newcommand{\colim}{\operatorname*{colim}}

% Arrows
\newcommand{\ra}{\rightarrow}
\newcommand{\from}{\leftarrow}
\newcommand{\To}{\Rightarrow}
\newcommand{\xto}{\xrightarrow}
\newcommand{\xfrom}{\xleftarrow}

\renewcommand{\labelitemi}{--} % Level 1: Dashes
\renewcommand{\labelitemii}{$\circ$} % Level 2: Circles
\renewcommand{\labelitemiii}{$\bullet$} % Level 3: Bullets (default)


\title{Notes: Category Theory in Context}
\author{Alyson Mei}
\date{\today}


\begin{document}

\maketitle

These notes are intended as a concise overview of the book \href{https://emilyriehl.github.io/files/context.pdf}{Category Theory in Context} by Emiliy Riehl.  

\tableofcontents

\section{Categories, functors, natural transformaions}

\subsection{Abstract and concrete categories}

\begin{definition}[Category] Notation: $\mc{C}, \mc{D}, \mc{E}, \ldots$
\end{definition}

\begin{remark}[Correspondence between objects and identity morphisms]
\end{remark}

\begin{example}[Concrete categories]
    $\mc{S}et$-based categories:
    \begin{itemize}
        \item $\mc{S}et$, $\mc{T}op$;
        \item $\mc{S}et_*$, $\mc{T}op_*$ (pointed);
        \item $\mc{G}rp$, $\mc{R}ing$, $\mc{F}ield$;
        \item $\mc{A}b\mc{G}rp$, $\mc{V}ect_{\mathbf{k}}$, $\mc{M}od_R$;
        \item $\mc{G}raph$, $\mc{D}ir\mc{G}raph$;
        \item $\mc{M}an$, $\mc{M}eas$ -- manifolds and measurable spaces;
        \item $\mc{P}oset$;
        \item $\mc{C}h_R$ -- chain complexes;
        \item $\mc{M}odel_\mathbb{T}$ -- $\sigma$-structures that model $\mathbb{T}$ (see universal algebra).
    \end{itemize}
\end{example}

\begin{remark*} 
    The category $\mc{S}et$ implicitly depends on the underlying set theory.
\end{remark*}

\begin{example}[Abstract categories]
    \hfill
    \begin{itemize}
        \item $\mc{M}at_R$ -- objects are integers, arrows $m \to n$ are $m \times n$ matrices with values in $R$;
        \item $\mathcal{B}G$;
        \item $(P, \leq)$ and ordinal $\alpha = \lbrace \beta \mid \beta < \alpha \rbrace$ as categories;
        \item $\mc{H}tpy$, $\mc{H}tpy_*$, $\mc{M}easure$.
    \end{itemize}
\end{example}

\begin{remark}[Russel's paradox]
\end{remark}

\begin{definition}[Small category]
    A category with only a set's worth of arrows.
\end{definition}

\begin{definition}[Locally small category] 
    A category with only set's worth of arrows between any pair of objects.
\end{definition}

\begin{definition}[Isomorphism, automorphism, endomorphism]
\end{definition}

\begin{example}[Isomorphisms in categories]
\end{example}

\begin{definition}[Groupoid] 
    A groupoid is a category in which every morphism is an isomorphism.
\end{definition}

\begin{example}[Examples of groupoids]
\end{example}

\begin{definition*}[Subcategory]
\end{definition*}

\begin{lemma}[Maximal groupoid] 
    Any category $\mc{C}$ contains a maximal groupoid, the subcategory generated by isomorphism of $\mc{C}$.
\end{lemma}

\begin{definition*}[Slice categories $c/\mc{C}$ and $\mc{C}/c$]    
\end{definition*}

\subsection{Duality}
\begin{definition}[Opposite category $\mc{C}^\op$]
\end{definition}

\begin{remark*}
    This duality is purely syntactic. Its usefulness becomes apparent when
    a) considering opposite constructions, b) working with functors.
\end{remark*}

\begin{example}[Opposite categories] 
    $\mc{M}at_R^\op$, $(P, \leq)^\op$, $(\mc{B}G)^\op$.
\end{example}

\begin{lemma}
    The following are equivalent (in a locally small $\mc{C}$):
    \begin{enumerate}
        \item $f: x \to y$ is an iso in $\mc{C}$.
        \item For all $c: \obj(\mc{C})$, post-composition with $f$ defines a bijection
        \[
            f_*: \mc{C}(c, x) \to \mc{C}(c, y).
        \]
        \item For all $c: \obj(\mc{C})$, pre-composition with $f$ defines a bijection
        \[
            f_*: \mc{C}(y, c) \to \mc{C}(x, c).
        \]
    \end{enumerate}
\end{lemma}

\begin{remark}[Note on representability]
\end{remark}

\addtocounter{theorem}{2}
\begin{definition}[Monomorphism, epimorphism]
    % https://q.uiver.app/#q=WzAsNCxbMSwwLCJ4Il0sWzIsMCwieSJdLFswLDAsInoiXSxbMywwLCJ3Il0sWzAsMSwiZiIsMCx7Im9mZnNldCI6LTF9XSxbMCwxLCJnIiwyLHsib2Zmc2V0IjoxfV0sWzIsMCwiXFx0ZXh0e2VwaX0iXSxbMSwzLCJcXHRleHR7bW9ub30iXV0=
    \[\begin{tikzcd}
        z & x & y & w
        \arrow["{\text{epi}}", from=1-1, to=1-2]
        \arrow["f", shift left, from=1-2, to=1-3]
        \arrow["g"', shift right, from=1-2, to=1-3]
        \arrow["{\text{mono}}", from=1-3, to=1-4]
    \end{tikzcd}\text{.}\]
    (to remember: epi $\to$ mono)
\end{definition}

\begin{example}[Mono in $\mc{S}et$ are precisely the injections]
\end{example}

\begin{example}[Section (right inverse), retraction (left inverse); split mono / epi]
\end{example}

\begin{example}[Iso $\Rightarrow$ mono and epi; counterexample $\mathbb{Z} \hookrightarrow \mathbb{Q}$ in $\mc{R}ing$ for the converse]
\end{example}

\begin{lemma}[Compositions of mono / epi]
    \hfill
    \begin{enumerate}
        \item Composable $f$ and $g$ are mono $\Rightarrow$ $gf$ is mono. Dually, $f$ and $g$ are epi $\Rightarrow$ $gf$ is epi.
        \item $gf$ is mono $\Rightarrow$ $f$ is mono. Dually, $gf$ is epi $\Rightarrow$ $g$ is epi.
    \end{enumerate}
\end{lemma}

\subsection{Functoriality}

\begin{definition}[Functor] Notation: $F: \mc{C} \to \mc{D}$.
\end{definition}

\begin{example}[Functors]
    \hfill
    \begin{itemize}
        \item Power set functor $P: \mc{S}et \to \mc{S}et$;
        \item Forgetful functors $U: \mc{C} \to \mc{S}et$, itermediate forgetful functors (e.g. $\mc{M}od_R \to \mc{A}b$, $\mc{G}r \to \mc{S}et_*$);
        \item Homotopy class functors $\mc{T}op_{(*)} \to \mc{H}tpy_{(*)}$;
        \item Fundamental group $\pi_1: \mc{T}op_* \to \mc{G}rp$ and its factor $\pi_1: \mc{H}tpy_* \to \mc{G}rp$ through $\mc{T}op_* \to \mc{H}tpy_*$, fundamental groupoid $\Pi_1: \mc{T}op \to \mc{G}roupoid$;
        \item $n$-cycles $Z_n$, $n$-boundary $B_n$ and $n$-th homology $H_n$ typed $\mc{C}h_R \to \mc{M}od_R$, all degrees simultaneously give $Z_*, B_*, H_*: \mc{C}h_R \to \mc{G}r\mc{M}od_R$ (codomain is the category of graded $R$-modules);
        \item Free functors $F: \mc{S}et \to \mc{C}$ (left adjoints to forgetful functors);
        \item Derivative $D: \mc{E}uclid_* \to \mc{M}at_\mathbb{R}$;
        \item $M^-: \mc{F}in_* \to \mc{S}et$ for a commutative monoid $M$.
    \end{itemize}
\end{example}

\begin{theorem}[Brouwer Fixed Point Theorem] Any continuous endomorphism of a 2-dimensional disk $D^2$ has a fixed point.
\end{theorem}

\begin{example}[Example for topological data analysis]
\end{example}

\begin{definition}[Contravariant functor] $F: \mc{C}^\op \to \mc{D}$.
\end{definition}

\addtocounter{theorem}{1}

\begin{example}[Contravariant functors]
    \hfill
    \begin{enumerate}
        \item Contravariant power set $P: \mc{S}et^\op \to \mc{S}et$;
        \item Dual vector space functor $*: \mc{V}ect_\mathbf{k}^\op \to \mc{V}ect_\mathbf{k}$;
    \end{enumerate}
\end{example}

\end{document}
